FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2024-08-01 paper.tex
jcon2: move future-work threads issue to the future work section

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\juliapkg}[1]{\texttt{#1.jl}}
\newcommand{\TrackedFloats}{\juliapkg{TrackedFloats}}
\newcommand{\TF}{\TrackedFloats}
\newcommand{\ShallowWaters}{\juliapkg{ShallowWaters}}
\newcommand{\OrdinaryDiffEq}{\juliapkg{OrdinaryDiffEq}}
\newcommand{\RxInfer}{\juliapkg{RxInfer}}
\newcommand{\Sherlogs}{\juliapkg{Sherlogs}}
\newcommand{\NBodySimulator}{\juliapkg{NBodySimulator}}
\newcommand{\Fp}{Floating-point}
\newcommand{\fp}{floating-point}
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zerocode}[1]{\zerowidth{\code{#1}}}
\newcommand{\genpropkill}{\emph{gen}-\emph{prop}-\emph{kill}}
\newcommand{\bigcheckmark}{\ding{51}}
% \newcommand{\bigxmark}{\ding{53}}
\newcommand{\tblnext}{\(\Rightarrow\)}
% \newcommand{\tblY}{\mbox{\bigcheckmark}}
% \newcommand{\tblN}{\scalebox{0.8}{\bigxmark}}
\newcommand{\tblY}{\checkmark}
\newcommand{\tblN}{\scalebox{1.2}{$\times$}}

@@

We note that \TF{} is currently not thread-safe; event logs are
collected globally and running \TF{} in a multithreaded environment
carries risks of corrupting the event log stack.
Improving \TF{} to be thread-safe is an area for future work.

@@

\subsection{\OrdinaryDiffEq{}}
\label{s:ode}

TODO awk.
With the targeted fuzzing abilities of \TF{}, we set focus on the
\NBodySimulator{} package from the SciML
team.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
We only wanted to test the robustness of the \NBodySimulator{} package, so we configured the fuzzer to inject \NaN{}s \emph{only} when inside of a function of that library.
This way, we could avoid injecting in the standard library or other dependencies of \NBodySimulator{} and concentrate on improving this one library.

However, Fuzzing discovered zero bugs and zero exceptional events.
In fact, fuzzing injected \emph{zero} exceptional values whatsoever, even with adjusting the odds to \emph{force} an injection at the first available opportunity.
The reason for this lack of \NaN{} injection, it turns out, is that the nbody
simulation merely creates a problem for the \OrdinaryDiffEq{}
solver.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}}
All the \fp{} operations were in another library.

Fuzzing on \OrdinaryDiffEq{} led to a curious situation.
The library itself reported a \NaN{} and printed a message that it would exit.
However, after printing that message, the program went into an infinite loop.
Using \CSTG{}s to guide our search, we found a \NaN{} kill that manifested repeatedly in the logs.
The stack traces for that kill originated from inside the file \texttt{solve.jl}:

@@

Finch is a domain-specific language for specifying
PDEs~\cite{heislerFinchDomainSpecific2022}.\footnote{Not to be confused
with the Finch loop optimizer~\cite{adka-cgo-2023}.}
In the spirit of FEniCS~\cite{fenics} and related
tools~\cite{freefem,openfoam,dune,firedrake},
Finch helps scientists quickly convert math into code.
What sets Finch apart is its flexibility.
It supports multiple discretization methods (finite element and finite
volume) and multiple backends (Julia, \CPP{}, \Dendro{}~\cite{dendro}).
Furthermore it strives to output code that humans can easily fine-tune.

@@

\subsection{Oceananigans}
\label{s:ocean}

\texttt{Oceananigans}~\cite{OceananigansJOSS} is simulation package for incompressible
fluid dynamics that can generate code for Nvidia GPUs.
For example, the following program (from the project readme) simulates turbulence:

@@

\subsection{Interface Concerns}

\TF{} is capable of monitoring all sorts of events and \fp{} values beyond exceptions.
For example, \TF{} could monitor large or small values with guidance from the user about which values are worth logging.
On a similar note, the interface to \TF{} is primitive: users express interest in a number by wrapping it
in a constructor such as \code{TrackedFloat64}.
Helper functions that track data structures and choose a default size would
make it easier to experiment with \TF{}.

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\juliapkg}[1]{\texttt{#1.jl}}
\newcommand{\TrackedFloats}{\juliapkg{TrackedFloats}}
\newcommand{\TF}{\TrackedFloats}
\newcommand{\ShallowWaters}{\juliapkg{ShallowWaters}}
\newcommand{\OrdinaryDiffEq}{\juliapkg{OrdinaryDiffEq}}
\newcommand{\RxInfer}{\juliapkg{RxInfer}}
\newcommand{\Sherlogs}{\juliapkg{Sherlogs}}
\newcommand{\NBodySimulator}{\juliapkg{NBodySimulator}}
\newcommand{\Oceananigans}{\juliapkg{Oceananigans}}
\newcommand{\Fp}{Floating-point}
\newcommand{\fp}{floating-point}
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zerocode}[1]{\zerowidth{\code{#1}}}
\newcommand{\genpropkill}{\emph{gen}-\emph{prop}-\emph{kill}}
\newcommand{\bigcheckmark}{\ding{51}}
% \newcommand{\bigxmark}{\ding{53}}
\newcommand{\tblnext}{\(\Rightarrow\)}
% \newcommand{\tblY}{\mbox{\bigcheckmark}}
% \newcommand{\tblN}{\scalebox{0.8}{\bigxmark}}
\newcommand{\tblY}{\checkmark}
\newcommand{\tblN}{\scalebox{1.2}{$\times$}}

@@

@@

\subsection{Fuzzing from NBody to ODE}
\label{s:ode}

We applied the fuzzing abilities of \TF{} to the
\NBodySimulator{} package.\footnote{\url{https://github.com/SciML/NBodySimulator.jl}}
To avoid injecting in the standard library or other dependencies,
we initially configured the fuzzer to inject \NaN{}s {only} when inside of a
function from the simulator.
Surprisingly, fuzzing injected \emph{zero} exceptional values even
when the fuzzer's odds were configured to force injection at the
first available opportunity.
It turns out that all the \fp{} operations for the simulation happened
within the \OrdinaryDiffEq{}
solver.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}}

Fuzzing on \OrdinaryDiffEq{} led to a curious situation.
The library itself reported a \NaN{} and printed a message stating that it would exit.
However, after printing that message, the program went into an infinite loop.
Using \CSTG{}s to guide our search, we found a \NaN{} kill that manifested repeatedly in the logs.
The stack traces for that kill originated from inside the file \texttt{solve.jl}:

@@

Finch is a domain-specific language for specifying
PDEs~\cite{heislerFinchDomainSpecific2022}.\footnote{Not to be confused
with the loop optimizer \juliapkg{Finch}~\cite{adka-cgo-2023}.}
In the spirit of FEniCS~\cite{fenics} and related
tools~\cite{freefem,openfoam,dune,firedrake},
Finch helps scientists quickly convert math into code.
What sets Finch apart is its flexibility.
It supports multiple discretization methods (finite element and finite
volume) and multiple backends (Julia, \CPP{}, \Dendro{}~\cite{dendro}).
Furthermore it strives to output code that humans can easily fine-tune.

@@

\subsection{\Oceananigans{}}
\label{s:ocean}

\Oceananigans{}~\cite{OceananigansJOSS} is simulation package for incompressible
fluid dynamics that can generate code for Nvidia GPUs.
For example, the following program (from the project readme) simulates turbulence:

@@

\subsection{Interface Concerns, Multi-Threading}

\TF{} could be extended to monitor events and values that go beyond exceptions,
such as very-large or very-small numbers according to bounds supplied by the user.
On a similar note, the interface to \TF{} is primitive: users express interest in a number by wrapping it
in a constructor such as \code{TrackedFloat64}.
Helper functions that track data structures and provide a default float size would
make experimentation easier.

\TF{} stores its configuration and event logs in global data structures,
making it unsafe for a multithreaded environment.
Adding locks would restore safety, but perhaps a richer interface is in order.
Individual threads may wish to configure local logging for better
data organization and performance.
